---
title: "R Notebook"
output: rmarkdown::github_document
---

---
title: "Bulk Data Processing -- TE analysis"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
Roman Reggiardo -- 11/16/2018  
Last Update -- 11/16/2018

### Setup:

```{r}
library(tidyverse)
library(ggpubr)
library(ggsci)
setwd('~/Desktop/Thesis/Projects/IPSC_KRAS/TEs/')
```

### Project Theme
```{r}
this_theme <- theme(text = element_text(size=15, face = 'bold', family = "sans"),
      legend.title = element_text(size = 20),
      axis.text.x = element_text(hjust=1, angle=40, color = 'black',
                                 size = 15, face = 'bold'),
      axis.text.y = element_text(color = 'black', 
                                 size=15, face = 'bold'),
      panel.background = element_rect(fill ='white'),
      axis.line.x = element_line(size=1),
      axis.line.y = element_line(size=1),
      strip.background = element_blank(),
      strip.text = element_text(face = 'bold', 
                                size = 5, hjust=0.95),
      legend.position='top',
      legend.key = element_blank(),
      axis.ticks.y = element_blank(), 
      axis.title.y = element_text(face='bold'),
      axis.ticks.x = element_line())
```

## Taking output from SalmonTE 

### 1. Parse into one data tidy set

Want the TPM, log2Fold Change and Pvalue of everything all in one spot  

*  Going to change naming convention slightly: ad for the adherenet d5 ipscs and eb for the d7 ebs.
```{r}
ad_stat <- read.csv('te_out/Stat_test/results.csv')
# filter by pvalue and keep only columns we are interested in
ad_stat <- ad_stat %>% filter(pvalue < 0.05)
trimm_ad_stat <- ad_stat[, c(1,2,3,5,8)]
ad_tpm <- read.csv('te_out/SalmonTE_output/EXPR.csv')
# rename columns to match experiment
names(ad_tpm) <- c('TE', 'SCRAM_1', 'KRAS_1', 'SCRAM_2', 'KRAS_2', 'SCRAM_3', 'KRAS_3')

####
eb_stat <- read.csv('eb_te_out/Stat_test/results.csv')
# filter by pvalue and keep only columns we are interested in
eb_stat <- eb_stat %>% filter(pvalue < 0.05)
trimm_eb_stat <- eb_stat[, c(1,2,3,5,8)]
eb_tpm <- read.csv('eb_te_out/SalmonTE_output/EXPR.csv')
# rename columns to match experiment
names(eb_tpm) <- c('TE', 'SCRAM_1', 'KRAS_1', 'SCRAM_2', 'KRAS_2', 'SCRAM_3', 'KRAS_3')

```

Need to figure out best merging structure and also filter by average TPM > 1:
```{r}
ad <- merge(trimm_ad_stat, ad_tpm, by.x = 'name', by.y = 'TE')
ad$tpMeans <- rowMeans(ad[, c(6:11)])
ad <- ad %>% filter(tpMeans > 1)
#ad$tpMeans <- NULL

####
eb <- merge(trimm_eb_stat, eb_tpm, by.x = 'name', by.y = 'TE')
eb$tpMeans <- rowMeans(eb[, c(6:11)])
eb <- eb %>% filter(tpMeans > 1)

```

Tidy:
```{r}
ad_tidy <- ad %>% gather(treatment, TPM, -c('name', 'class', 'clade', 
                                 'log2FoldChange', 'pvalue', 'tpMeans')) %>%
  separate(treatment, sep = '_', 1, into = c('Treatment', 'Rep')) %>%
  group_by(name)

eb_tidy <- eb %>% gather(treatment, TPM, -c('name', 'class', 'clade', 
                                 'log2FoldChange', 'pvalue', 'tpMeans')) %>%
  separate(treatment, sep = '_', 1, into = c('Treatment', 'Rep')) %>%
  group_by(name)

```

Plot:
```{r}
ad_volc <- 
  ggplot(ad, aes(log2FoldChange, -log10(pvalue), color = clade)) + 
  this_theme + 
  geom_point() +
  geom_vline(xintercept = 0, linetype='dotted') +
  ggrepel::geom_text_repel(aes(label = ifelse(tpMeans > 1, 
                       as.character(name), '')),
                           hjust = 0,
                           direction = 'y', 
                       color= 'black',
                       segment.size = 0,
                       point.padding = 1.25) + 
  scale_color_npg() +
  scale_x_continuous(limits = c(-5,5))

eb_volc <- 
  ggplot(eb, aes(log2FoldChange, -log10(pvalue), color = clade)) + 
  this_theme + 
  geom_point() +
  geom_vline(xintercept = 0, linetype='dotted') +
  ggrepel::geom_text_repel(aes(label = ifelse(tpMeans > 1, 
                       as.character(name), '')),
                           hjust = 0,
                           direction = 'y', 
                       color= 'black',
                       segment.size = 0,
                       point.padding = 1.25) + 
  scale_color_npg() +
  scale_x_continuous(limits = c(-5,5))

eb_map <- ggplot(eb, aes(clade, name, 
                         fill=log2FoldChange, 
                         label = round(tpMeans))) +
  geom_tile() +
  geom_text() +
  this_theme +
  scale_fill_gsea()+ 
  ylab('') + xlab('')

ad_map <- ggplot(ad, aes(clade, name, 
                         fill=log2FoldChange, 
                         label = round(tpMeans))) +
  geom_tile() +
  geom_text() +
  this_theme +
  scale_fill_gsea() + 
  ylab('') + xlab('')

eb_volc
eb_map
ad_map
ad_volc
```

Compare expression of DE TEs:
```{r}
ad$which <- 'ad'
eb$which <- 'eb'
eb <- eb %>% filter(name %in% ad$name)
ad <- ad %>% filter(name %in% eb$name)
ad_eb <- rbind(ad, eb)

comp_te <- ggplot(ad_eb, aes(name, log2FoldChange, fill=which)) + 
  geom_bar(stat = 'identity', position = 'dodge') +
  this_theme +
  theme(axis.text.x = element_text(hjust=1, angle=40)) +
  xlab('') + 
  scale_fill_manual(labels = c('iPSC','EB'), 
                    values = c('purple', 'forestgreen'),
                    name='') + 
  geom_hline(yintercept = 0, linetype='dotted')
```

Now open the bed intersection from long read data:
```{r}
intersect_one <- read.delim('parsed_int_01.txt', header = F)
head(intersect_one)


```

Before moving on want to look at read/alignment lengths:  
*lines are placed at 1kb, 1.55kbm 2.45kb*
```{r}
repOneStats <- read.delim('read_stats.txt', sep = ' ', header = F)
ggplot(repOneStats) + 
  geom_density(aes(V1, V2), stat = 'identity') +
  this_theme + ylab('') + xlab('') + 
  geom_vline(xintercept = 1000) + 
  geom_vline(xintercept = 1550) + 
  geom_vline(xintercept = 2450) +
  ggtitle('Read Length distribution of Rep-aligned subset')

alnRepOneStats <- read.delim('unfiltered_aln_length.txt', header = F) 
ggplot(alnRepOneStats) + 
  geom_density(aes(V1)) + 
  this_theme + ylab('') + xlab('') + 
  geom_vline(xintercept = 1000) + 
  geom_vline(xintercept = 1550) + 
  geom_vline(xintercept = 2450)+
  ggtitle('Read Length distribution of Alignments')

filtAlnRepOneStats <- read.delim('filtered_aln_length.txt', header = F)
ggplot(filtAlnRepOneStats) + 
  geom_density(aes(V1)) + 
  this_theme + ylab('') + xlab('') + 
  geom_vline(xintercept = 1000) + 
  geom_vline(xintercept = 1550) + 
  geom_vline(xintercept = 2450)+
  ggtitle('Read Length distribution of Alignments filtered by length')

filtAlnRepOneQual <- read.delim('filtered_qual_spread.txt', header = F)
ggplot(filtAlnRepOneQual) + 
  geom_density(aes(V1)) + 
  this_theme + ylab('') + xlab('') + 
  geom_vline(xintercept = 3) +
  geom_vline(xintercept = 30)+
  ggtitle('MAPQ Score distribution of Rep-aligned subset')

full_filt_len <- read.delim('full_filt_aln_length.txt', header = F)
ggplot(full_filt_len) + 
  geom_density(aes(V1)) + 
  this_theme + ylab('') + xlab('') + 
  geom_vline(xintercept = 1000) + 
  geom_vline(xintercept = 1550) + 
  geom_vline(xintercept = 2450)+
  ggtitle('Read Length distribution of quality and length filtered Alignments')
```

Load in the bed created from the filtered alignments:
```{r}
read_ref <- read.delim('read_to_repeat_ref.txt', header=F)
read_alnLen <- read.delim('fixed_read_alnLen.txt', header = F, sep = '')
filt_intersect_one <- read.delim('parsed_filt_int_01.txt', header = F)
nosplit_intersect_one <- read.delim('nosplit_filt_01_intersect.bed', header = F)
te_only_one <- read.delim('repeat_only_intersect.bed', header = F)
read_len <- read.delim('read_length.txt', header = F)
names(read_len)[2] <- 'aln_len'

te_only_one_length <- merge(te_only_one, read_len, by.x = 'V4', by.y = 'V1')
te_only_one_length <- te_only_one_length %>% filter(V11/aln_len > 0.50)
te_only_one_length$aln_len <- te_only_one_length$V3 - te_only_one_length$V2.x
te_only_one_length <- te_only_one_length %>% group_by(V1, V10)
te_only_one_length %>% count(V1, V10)

ref <- merge(read_alnLen, read_ref, by.x = 'V1', by.y = 'V1')
ref <- ref %>% separate(V2.y, sep='_',1, into=c('g','r','TE'))
ref$g <- NULL
ref$r <- NULL
ref <- ref %>% unique()
ref <- drop_na(ref)
ref <- ref %>% 
  filter(V1 %in% filt_intersect_one$V4)
ref <- ref %>% filter(!grepl(')n', TE))
ref <- ref %>% filter(!grepl('-rich', TE))
names(ref) <- c('READ','ALN_LEN','TE')
ref <- ref %>% filter(ALN_LEN > 1)

te_only_ref 
merge(ref, te_only_one, by.x = 'READ', by.y = 'V4') %>% unique()

intersect_ref <- merge(filt_intersect_one, ref, by.x = 'V4', by.y = 'READ')

overlap_filter_int_ref <- intersect_ref %>% 
  filter(V9/ALN_LEN > 0.9) %>% 
  filter(ALN_LEN > 1) %>%
  unique()

overlap_filter_int_ref <- overlap_filter_int_ref %>% filter(V9/ALN_LEN < 1.2)
overlap_filter_int_ref <- overlap_filter_int_ref %>% filter(V1 != 'chrM')

length(unique(overlap_filter_int_ref$V4))

lnc_int_ref <- overlap_filter_int_ref %>% 
  filter(V8 %in% consensus_tpm$GENE.x)

gene_table <- as.data.frame(table(lnc_int_ref$V8))

te_table <- as.data.frame(table(te_only_one$V10))


tePLot <- ggplot(subset(te_table, Freq > 20)) +
  geom_bar(aes(reorder(Var1, Freq), Freq), stat = 'identity') + 
  this_theme + coord_flip()
gePlot <- ggplot(subset(gene_table, Freq > 1)) +
  geom_bar(aes(reorder(Var1, Freq), Freq), stat = 'identity') +
  this_theme + coord_flip()

ggplot(subset(te_only_one)) +
  geom_bar(aes(V7)) + this_theme

ggplot(nosplit_intersect_one) +
  geom_bar(aes(V9)) + this_theme

chr_overview <- ggplot(te_only_one_length) +
  geom_bar(aes(V1, fill=V10)) + 
  this_theme + scale_fill_discrete()

chr3 <- te_only_one_length %>% subset(V1 == 'chr3') 

chr3 %>% group_by_('V2.x', 'V10') %>% summarise(freq = ..count..)

chr3 <- chr3 %>% count(V2.x, V10, V6)


chr3_overview <- ggplot(chr3) + 
  geom_bar(aes(V2.x, 
               ifelse(V6 == '-', -1 * n, n),
               color = V10),
           stat='identity') + 
  this_theme

```


No overlap with Gencode alignments, annotated with repeat masker:
```{r}
no_ol <- read.delim('no_ol.aln_01_sort.q45.rep_int.bed', header=F)

no_ol <- no_ol %>% separate(V21, sep = ';', 1, into = c('g', 't', 'f', 'c'))
no_ol$g <- substring(no_ol$g, 9)
no_ol$c <- substring(no_ol$c, 11)
no_ol <- subset(no_ol, grepl('_', V1) != T)
length(unique(no_ol$V4))

no_ol_chr <- unique(no_ol[,c(1,2,4,24)])

ltr7_hervh <- subset(no_ol, g %in% c('LTR7', 'HERVH-int'))
ltr7 <- unique(subset(ltr7_hervh, g == 'LTR7')[,c(1,2,4)])
names(ltr7)[]
HERVH <- unique(subset(ltr7_hervh, g == 'HERVH-int')[,c(1,2,4)])
ltr_erv_dist <- ggplot(ltr7_hervh) + 
  geom_density(aes(V1, fill=g), position = 'dodge') +
  facet_wrap(~V1)  +
  this_theme + 
  theme(axis.text.x = element_blank(),
                          axis.ticks.x = element_blank()) + 
  scale_fill_aaas(name = '', alpha = 0.4) + 
  xlab('') + ylab('')

te_alone <- ggplot(no_ol) + 
  geom_bar(aes(V1, fill=c)) + 
  this_theme + 
  scale_x_discrete(limits = c('chr1', 'chr2', 'chr3', 'chr4', 'chr5',
                              'chr6', 'chr7', 'chr8', 'chr9',
                              'chr10', 'chr11', 'chr12', 'chr13',
                              'chr14', 'chr15', 'chr16', 'chr17',
                              'chr18', 'chr19', 'chr20', 'chr21',
                              'chr22', 'chrX', 'chrY')) + 
  scale_fill_ucscgb(name='') + xlab('')


consensus_up <- read.csv('~/Desktop/Thesis/Projects/IPSC_KRAS/consensus_up.csv')
# cut -d',' -f3 consensus_up.csv | tr -d '""' | sort | uniq > consensus_up_enst.txt
# go to UCSC_GB to get coordinates
consensus_tss <- read.delim('~/Desktop/Thesis/Projects/IPSC_KRAS/consensus_up_chr_tss.txt')
consensus_tss <- consensus_tss %>% mutate(TSS = ifelse(strand == '+', txStart, txEnd))
consensus_up_merge <- merge(consensus_tss, consensus_up, by.x = 'X.name', by.y = 'ENST')

consensus_up_merge <- consensus_up_merge %>% dplyr::group_by('chrom')

# https://stackoverflow.com/questions/6988184/combining-two-data-frames-of-different-lengths
cbindPad <- function(...){
args <- list(...)
n <- sapply(args,nrow)
mx <- max(n)
pad <- function(x, mx){
    if (nrow(x) < mx){
        nms <- colnames(x)
        padTemp <- matrix(NA, mx - nrow(x), ncol(x))
        colnames(padTemp) <- nms
        if (ncol(x)==0) {
          return(padTemp)
        } else {
        return(rbind(x,padTemp))
          }
    }
    else{
        return(x)
    }
}
rs <- lapply(args,pad,mx)
return(do.call(cbind,rs))
}

bind_ltr_up <- cbindPad(consensus_up_merge, ltr7)

chr20_bind <- cbindPad(subset(consensus_up_merge, chrom == 'chr20'), 
         subset(ltr7, V1 == 'chr20'))

consensus_up_trim <- consensus_up_merge[, c(1,2,6)]
ltr7 <- add_column(ltr7, X.name = 'ltr7', .before = 1)
consensus_up_trim <- add_column(consensus_up_trim, read = '1', .after = 3)
consensus_up_trim$X.name <- 'ENST'
names(ltr7)[c(2,3,4)] <- c('chrom', 'TSS', 'read') 
bind.trim.ltr7 <- rbind(ltr7, consensus_up_trim)

ggplot() + 
  geom_density(aes(x=TSS), data = consensus_up_trim, fill = 'green', alpha=.3) +
  geom_density(aes(x=TSS), data = ltr7, fill = 'yellow', alpha=.3) + 
  this_theme + 
  facet_wrap(~chrom, nrow=3, ncol = 8, scales = 'free') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

ggplot(chr20_bind) + 
  geom_point(aes(TSS, V2)) + 
  this_theme 
```

Overlap:
```{r}
ol <- read.delim('ol.aln_01_sort.q45.rep_int.bed', header=F)
ol <- ol %>% separate(V9, sep = ';', 1, into = c('1','2','3','4','5','6'))
ol <- unique(ol[, c(1,2,3,4,5,6,8,10,11,12)])
ol <- ol %>% separate(V30, sep = ';', 1, into = c('g', 't', 'f', 'c'))
ol$g <- substring(ol$g, 9)
ol$c <- substring(ol$c, 11)
ol <- subset(ol, grepl('_', V1) != T)
ol <- subset(ol, !(c %in% c('Satellite', 'RNA', 'RC?')))
length(unique(subset(ol, grepl('gene', `4`) == T)$`6`))

ol_chr <- unique(ol[, c(1,6,7,8,9,13)])
ol_chr <- subset(ol_chr, grepl('gene', `4`) == T)
ol_chr <- unique(ol_chr[, c(1,3,5,6)])
ol_genes <- substring(unique(ol_chr$`4`), 11)
print(unlist(ol_genes))
te_derived <- ggplot(ol) + 
  geom_bar(aes(V1, fill=c)) + 
  this_theme + 
  scale_x_discrete(limits = c('chr1', 'chr2', 'chr3', 'chr4', 'chr5',
                              'chr6', 'chr7', 'chr8', 'chr9',
                              'chr10', 'chr11', 'chr12', 'chr13',
                              'chr14', 'chr15', 'chr16', 'chr17',
                              'chr18', 'chr19', 'chr20', 'chr21',
                              'chr22', 'chr23', 'chrX', 'chrY')) + 
  scale_fill_ucscgb(name='') + xlab('')

ol
```

Compare with lncs from https://www.nature.com/articles/nature13804/figures/8:
```{r}
library(readxl)
hervLncs <- read_xlsx('~/Downloads/nature13804-s2/Supplementary Table 10.xlsx')
hervCons <- subset(consensus, GENE %in% hervLncs$gene_name)

merge(hervCons, hervLncs, by.x = 'GENE', by.y = 'gene_name' )
```


